---
title: "Insect Duration Models"
author: "Michael W Belitz"
date: "5/13/2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(lme4)
library(lmerTest)
library(dplyr)
library(effects)
library(ggplot2)
library(car)
```

## Number of species with at least 5 comparable cells per Order

```{r, fivecells, echo = FALSE, message=FALSE}
model_df <- read.csv(file = "data/model_dfs/duration_climate_population_data.csv",
                     stringsAsFactors = FALSE) 

id_cells <- model_df %>% 
  group_by(lon, lat) %>% 
  summarise(count = n()) %>% 
  tibble::rownames_to_column() %>% 
  rename(id_cells = rowname)

model_df <- left_join(model_df, id_cells)

model_df2 <- model_df %>% 
  na.omit() %>% 
  mutate(temp = scale(temp),
         prec = scale(prec),
         pop = scale(log10(pop)),
         prec_seas = scale(bio15),
         temp_seas = scale(bio4),
         onsetci = onset_high - onset_low,
         offsetci = offset_high - offset_low)

datadens <- model_df2 %>% 
  group_by(scientificName, Order) %>% 
  summarise(count = n())

has_5_cells <- filter(datadens, count >= 5) %>% 
  filter (scientificName != "Apis mellifera")#300 species 

ggplot(has_5_cells, aes(x = Order)) + 
  geom_bar(stat = "count", aes (fill = Order)) +
  labs(y = "Number of Species") +
  ggtitle("At Least 5 Comparable Cells - 300 total spp") +
  theme_bw()
```

# Number of species with at least 10  comparable  cells per Order

Because the proportion of number of Leps per all other orders didn't increase with having at least 10 comparable cells, I decided to only use species with at least 10 comparable cells for the analyses. 

```{r, tencells, echo=FALSE, message=FALSE}

has_10_cells <- filter(datadens, count >= 10) %>% 
  filter (scientificName != "Apis mellifera") # 217

ggplot(has_10_cells, aes(x = Order)) + 
  geom_bar(stat = "count", aes (fill = Order)) +
  labs(y = "Number of Species") +
  ggtitle("At Least 10 Comparable Cells - 217 total spp") +
  theme_bw()

```


# Model Results

## Onset
Top model
```{r onset, echo=TRUE}
m9on <- lmer(onset ~ temp + pop + prec + prec_seas + temp_seas + temp:pop +
               (1|id_cells) + (1|scientificName) +
               (0 + temp | scientificName) + 
               (0 + pop | scientificName) + 
               (0 + prec | scientificName) +
               (0 + temp_seas | scientificName) + 
               (0 + prec_seas | scientificName) + 
               (0 + temp:pop | scientificName),
             data = model_df2, REML = FALSE, 
             lmerControl(optimizer = "bobyqa"))

summary(m9on) 

plot(effects::effect("temp:pop", m9on), multiline = T)
MuMIn::r.squaredGLMM(m9on)
car::vif(m9on)
class(m9on) = "lmerMod"
performance::check_collinearity(m9on)
performance::model_performance(m9on)
```

## Offset
Top model
```{r offset, echo=TRUE}
m7off <- lmer(offset ~ temp + pop + prec + prec_seas + temp_seas + temp:pop +
                (1|id_cells) + (1|scientificName) +
                (0 + temp | scientificName) + 
                (0 + pop | scientificName) + 
                (0 + prec | scientificName) +
                (0 + temp_seas | scientificName) + 
                (0 + prec_seas | scientificName) +
                (0 + temp:pop | scientificName),
              data = model_df2, REML = FALSE, 
              lmerControl(optimizer = "bobyqa"))

summary(m7off) 

plot(effects::effect("temp:pop", m7off), multiline = T)
MuMIn::r.squaredGLMM(m7off)
car::vif(m7off)
class(m7off) = "lmerMod"
performance::check_collinearity(m7off)
performance::model_performance(m7off)
```

## Duration
Top model
```{r duration, echo=TRUE}
m6dur <- lmer(duration ~ temp + pop + temp_seas + temp:pop + prec_seas + prec +
                (1|id_cells) + (1|scientificName) +
                (0 + temp | scientificName) + 
                (0 + pop | scientificName) +
                (0 + prec_seas | scientificName) +
                (0 + temp:pop | scientificName) +
                (0 + temp_seas | scientificName) +
                (0 + prec | scientificName),
              data = model_df2, REML = FALSE, 
              lmerControl(optimizer = "bobyqa"))

summary(m6dur) 
plot(effects::effect("temp:pop", m6dur), multiline = T)
MuMIn::r.squaredGLMM(m6dur)
car::vif(m6dur)
performance::check_collinearity(m6dur)
performance::model_performance(m6dur)
```
