---
title: "Data Density of Potential Insect Species"
author: "Michael W Belitz"
date: "2/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width = 12, fig.height = 8)
```

```{r mappingScript, echo=FALSE, warning=FALSE, message=FALSE}
library(rgbif)
library(dplyr)
library(lubridate)
library(ggplot2)
library(rnaturalearth)
library(sf)
library(ridigbio)

# function to map gbif occurrence data

heat_map_obs <- function(binomial){
  
  idig <- idig_search_records(limit = 10000, rq = list(scientificname = binomial), 
            fields = c("data.dwc:decimalLongitude", "data.dwc:decimalLatitude",
                       "data.dwc:year", "data.dwc:month", "data.dwc:eventDate",
                       "data.dwc:scientificName", "data.dwc:basisOfRecord")) 
  
  names(idig) <- c("decimalLongitude", "decimalLatitude", "year", "month", "eventDate",
                    "scientificName", "basisOfRecord")
  
  #rgbif
gbif <- occ_search(scientificName = binomial,
                      hasCoordinate = TRUE, country = "US",
                      limit = 200000)

gbif_df <- gbif$data %>% 
  filter(basisOfRecord == "HUMAN_OBSERVATION") %>% 
  select(decimalLongitude, decimalLatitude, year, month, eventDate, 
         scientificName, basisOfRecord)
  
total <- rbind(gbif_df, idig) %>% 
  filter(!is.na(year)) %>% 
  filter(!is.na(decimalLatitude)) %>% 
  filter(!is.na(decimalLongitude)) %>% 
  filter(!is.na(eventDate))
  
total$decimalLongitude <- as.numeric(total$decimalLongitude)
total$decimalLatitude <- as.numeric(total$decimalLatitude)
total$year <- as.numeric(total$year)

total <- total %>% 
  mutate(decade = floor(year/10)*10)

  us <- rnaturalearth::ne_countries(country = "United States of America", 
                                 returnclass = "sf")
  
ggplot() + 
  geom_sf(us, mapping = aes()) +
  geom_bin2d(total,mapping = aes(x = decimalLongitude, y = decimalLatitude), alpha = 0.9, binwidth = c(2,2)) + 
  coord_sf( xlim = c(-125, -65), ylim = c(20, 60)) +
  scale_fill_continuous(type = "viridis",trans = "log10") + 
  facet_wrap(~decade)
}

## Function to make summary table

summarize_occ_data <- function(binomial){
  dat <- occ_search(scientificName = binomial, hasCoordinate = TRUE, country = "US", limit = 200000)
  
  sp_df <- dat$data %>% 
    mutate(doy = lubridate::day(eventDate)) %>% 
    filter(!is.na(year)) %>% 
    filter(!is.na(doy))
  
  sum_table <- sp_df %>% 
    group_by(year, basisOfRecord) %>% 
    summarise(count = n())
  
  sum_table <- sum_table %>% 
    filter(count >= 10)
  
  return(sum_table)
}
```


## Data density of potential insects 

For this second exploration, I wrote a function that would map the observations by decades as a heat map.

## Saturniidae

Automeris io

```{r, io}
heat_map_obs("Automeris io")
```

Eacles imperialis

```{r, eacles, cache=TRUE}
heat_map_obs("Eacles imperialis")
```

Antheraea polyphemus

```{r, polyphemus, cache=TRUE}
heat_map_obs("Antheraea polyphemus")
```

Dryocampa rubicunda

```{r, dryocampa, cache=TRUE}
heat_map_obs("Dryocampa rubicunda")
```

Hemileuca eglanterina

```{r, hemileuca, cache=TRUE}
heat_map_obs("Hemileuca eglanterina")
```


## Cicadidae

Platypedia putnami

```{r, platypedia,cache=TRUE}
heat_map_obs("Platypedia putnami")
```

Magicicada septendecim

```{r, septendecim, cache=TRUE}
heat_map_obs("Magicicada septendecim")
```

Tibicen esfandiarii
```{r, esfandiarii, cache=TRUE}
heat_map_obs("Tibicen esfandiarii")
```

## Ephemeroptera

Stenacron interpunctatum
```{r, stenacron,cache=TRUE}
heat_map_obs("Stenacron interpunctatum")
```

## Odonata

Ischnura verticalis

```{r, Ischnura, cache=TRUE}
heat_map_obs("Ischnura verticalis")
```

## Megaloptera

Nigronia serricornis

```{r, Nigronia serricornis, cache=TRUE}
heat_map_obs("Nigronia serricornis")
```

## Silphidae 

Nicrophorus orbicollis

```{r, Nicrophorus orbicollis, cache=TRUE}
heat_map_obs("Nicrophorus orbicollis")
```

## Coccinellidae

```{r, Hippodamia, cache=TRUE}
heat_map_obs("Hippodamia convergens")
```